### 2018-05-12 1_intro.tex
Finish up to proc units.

\item Some other tools perform static instrumentation of concurrent C programs and encode the WMM into the program representation so that it can be model-checked by standard tools.
The examples are
the instrumenting compiler \tool{goto-cc} which is a part of \tool{CBMC} model checker~\cite{kroening2014cbmc},
the program transformation tool~\cite{alglave2013software},
the tool \tool{Weak2SC} for producing program descriptions which can be fed into standard model checking tools (such as \tool{SPIN}~\cite{holzmann1997model} or \tool{NuSMV}~\cite{cimatti2000nusmv}) for performing memory model-aware analysis~\cite{travkin2016verification}.
\end{itemize}

>>>>

\item Some other tools perform static instrumentation of concurrent C programs and encode the WMM into the program representation so that it can be model-checked by standard tools.
The examples are
%TODO: cite the 'wmm-general/Thread-Modular Static Analysis for Relaxed Memory Models.pdf'
the instrumenting compiler \tool{goto-cc} which is a part of \tool{CBMC} model checker~\cite{kroening2014cbmc},
the tool that performs the sequentialisation of concurrent programs~\cite{alglave2013software},
the tool \tool{Weak2SC} for producing program descriptions which can be fed into standard model checking tools (such as \tool{SPIN}~\cite{holzmann1997model} or \tool{NuSMV}~\cite{cimatti2000nusmv}) for performing memory model-aware analysis~\cite{travkin2016verification}.
\end{itemize}

### 2018-05-12 4_impl.tex
Finish up to proc units.

  \1 \texttt{XControlFlowEvent} \\
    The control-flow event indicates a non-linear jump in the code. We distinguish two kinds of control-flow events: 
    %\begin{itemize}[noitemsep]
      \2 \texttt{XJumpEvent} that performs no computation and no data operation, it can be safely removed from the model as an optimisation,
      \2 \texttt{XMethodCallEvent} that models the function call with the \texttt{fastcall} calling convention (passing arguments in registers).
      %TODO: check the diagrams if the XMCE really inherits XLMU
      The function call also implements the \texttt{XLocalMemoryUnit} since it represents the computed value returned by the function call.
      If the function has been resolved, the actual arguments are bind to the formal parameters (treated as temporary local memory units), the called \texttt{XMethodCallEvent} is pushed onto the call stack, and the execution jumps to the function body.
      Each \texttt{return} statement creates the assignment of the function call event on the top of call stack.
      \\
      If the function has not been resolved, \textbf{TODO: think about it. Will the formula be SAT? Or do we need to abort the analysis?}
    %\end{itemize}

@@

  \1 \texttt{XFakeEvent} \\
    The fake events are auxiliary elements of \xgraph{}.
    %\begin{itemize}[noitemsep]
    \2 \texttt{XEntryEvent}, the per-process unique source event in the event-flow graph,
    \2 \texttt{XExitEvent}, the process sink event. There are two kinds of sink events: \textit{complete} and \textit{incomplete} (see description in Section~\ref{ch:impl:proc:x-unroll}),
    \2 \texttt{XNopEvent}, the no-operation event (a jump to the next event), used for correct encoding in case when the control-flow branch does not have any event (see Figure~\ref{encode:branching:nop}).
   % \end{itemize}
%\end{itemize}
\end{outline}

@@

\begin{enumerate}[noitemsep]
\item the graph must have a single source with no ingoing edges, and two sinks of different kinds without outgoing edges,
\item the graph must be connected,
\item each node of the graph can have either one or two direct control-flow successors,
\item only nodes of type \texttt{XComputationEvent} can have two direct control-flow successors,
%\item each node except the source node must have at least one parent
\item a \co-edge connects two writes, an \rf-relation edge connects a write and a read,
\item all write-event nodes except initial write-nodes must have exactly one \co-predecessor
\item all write-event nodes except final write-nodes must have exactly one \co-successor.
%TODO: \item \textbf{TODO: more?}
\end{enumerate}

@@

A \wmodel{} represents a recursively defined AST of relations defined by the memory model and an assertion expression.
The atomic elements of \wmodel{} are the basic relations (\po, \rf and \co; see Section~\ref{ch:wmm:model:relations}) and sets of events ($\mathbb{R}$, $\mathbb{W}$, $\mathbb{IW}$, etc.; see Section~\ref{ch:wmm:model:events}).
The expressions of \wmodel{} are unary (such as complement, transitive closure, etc.) and binary operations (such as union, intersection, etc.) over relations or sets of events; see Section~\ref{ch:wmm:cat}.

<todo: enumerate elements>

<carries the origin -- from cat file>

@@

Following the structure of a Z3-formula, the atomic elements of a \zformula{} are variables, and the expressions are

reference to Z3-tutorial AND

All expressions and variables of a \zformula{} are typed.
Basically, we distinguish \textit{boolean} and \textit{numeric} (integer, floating point, <else?>) data type

%typed: Expr, BoolExpr, ArithExpr, ...

carries the origin (X)

<todo: enumerate elements>

@@

Construction of immutable data structures is performed by units that follow the \textit{builder} pattern.
<TODO: SAY MORE: validation of ht einput elements during construction, maybe verif after building>

@@

The input memory model language \cat{} is discussed in Section~\ref{ch:wmm:cat}.
The \tool{ANTLR} grammar for the \cat{} language was extracted from the parser used by \tool{herd} tool%
%
\footnote{\tool{herd} project repository: \url{https://github.com/herd/herdtools7}} %
%
written in OCaml.
%TODO: upload the grammar to github
%The input memory model 
\textbf{TODO: add reference to the grammar file in github repository OR include a short version as an appendix}

@@

%<not implemented yet>

@@

The \xgraph{} abstract machine model disposes infinitely many memory units, both local and shared (see Section~\ref{ch:impl:model:xgraph:mem}).
The X-compiler accesses all memory units via the \texttt{XMemoryManager}, which by the end of pre-compilation stage is already initialised (has registered all shared memory units).
However, the memory manager is a stateful component of the compiler as it offers the methods for declaring and removing local memory units dynamically at the compilation time.
At each time of the compilation process, the \texttt{XMemoryManager} can resolve the memory unit by its name.
Following the C standard, local memory units have higher priority over global ones.
Since the C language allows the variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} carries the block-context stack. %TODO: refactor XMemoryManager!

@@

The invocation hooking module serves as a knowledge base that stores the semantics of functions.
%TODO: rename HookManager -> XHookManager
Once the X-compiler meets the function invocation, it calls the \texttt{XHookManager}, which tries to match the function signature (in case of program in C language -- only the function name) across all signatures it stores.
Once the signature matches, the hook manager intercepts the compilation process and interprets the method according to the algorithm it stores.
%TODO: say sth about the mode
The hook manager operates multiple invocation hooks depending on the user-defined mode (
 \texttt{XInvocationHook}

%TODO: rename InvocationHook -> XInvocationHook
multiple hooks for mult modes
%, having access to the X-interpreter (see Section~\ref{ch:impl:proc:x-compiler:compilation}),

>>>>

  \1 \texttt{XControlFlowEvent} \\
    The control-flow event indicates a non-linear jump in the code. We distinguish two kinds of control-flow events: 
    \2 \texttt{XJumpEvent} that performs no computation and no data operation, it can be safely removed from the model as an optimisation,
    \2 \texttt{XMethodCallEvent} that models the function call with the \texttt{fastcall} calling convention (passing arguments in registers).
    %TODO: check the diagrams if the XMCE really inherits XLMU
    The function call also implements the \texttt{XLocalMemoryUnit} since it represents the computed value returned by the function call.
    If the function has been resolved, the actual arguments are bind to the formal parameters (treated as temporary local memory units), the called \texttt{XMethodCallEvent} is pushed onto the call stack, and the execution jumps to the function body.
    The call stack is bounded by the user-defined parameter; once the stack is full, the interpretation continues without jumping to the body of the invoked function (such cases are properly logged). %TODO: call-stack: user-defined parameter
    Each \texttt{return} statement creates the assignment of the function call event on the top of call stack.
    Note, this approach works for any kind of recursion, which is unrolled up the user-defined boundary as well as explicit loops.
    \\
    According to the Rice's theorem~\cite{rice1953classes}, in general case there always will be functions unresolved within an analysis launch.
    If the function has not been resolved, it can be safely assumed to be a no-operation function.
    In other words, we suppose that the knowledge base of \porthos{} is complete and the tool can resolve all memory-operation functions and fence instructions.
    Although this can affect the completeness of the analysis, it is safe to make such an assumption as the user can check the log file and manually analyse the semantics of an unresolved call and it to the knowledge base if necessary.

@@

  \1 \texttt{XFakeEvent} \\
    The fake events are the auxiliary elements of \xgraph{}.
    %\begin{itemize}[noitemsep]
    \2 \texttt{XEntryEvent}, the per-process unique source event in the event-flow graph,
    \2 \texttt{XExitEvent}, the process sink event. There are two kinds of sink events: \textit{complete} and \textit{incomplete} (see description in Section~\ref{ch:impl:proc:x-unroll}),
    \2 \texttt{XNopEvent}, the no-operation event (a jump to the next event), used for correct encoding in case when the control-flow branch does not have any event (see Figure~\ref{encode:branching:nop}).
   % \end{itemize}
%\end{itemize}
\end{outline}

@@

\begin{enumerate}[noitemsep]
\item the graph must have a single source with no ingoing edges, and two sinks of different kinds without outgoing edges,
\item the graph must be connected,
\item each node of the graph can have either one or two direct control-flow successors,
\item only nodes of type \texttt{XComputationEvent} can have two direct control-flow successors,
%\item each node except the source node must have at least one parent
\item a \co{}-edge connects two writes, an \rf{}-relation edge connects a write and a read,
\item all write-event nodes except initial write-nodes must have exactly one \co-predecessor
\item all write-event nodes except final write-nodes must have exactly one \co-successor.
%TODO: \item \textbf{TODO: more?}
\end{enumerate}

@@

A \wmodel{} represents a recursively defined AST of relations defined by the memory model and an assertion expression.
The atomic elements of \wmodel{} are the basic relations (\po, \rf and \co; see Section~\ref{ch:wmm:model:relations}) and sets of events ($\mathbb{R}$, $\mathbb{W}$, $\mathbb{IW}$, etc.; see Section~\ref{ch:wmm:model:events}).
The expressions of \wmodel{} are unary (such as complement, transitive closure, etc.) and binary operations (such as union, intersection, etc.) over relations or sets of events; see Section~\ref{ch:wmm:cat}.
For the sake of transparency, each element of a \wmodel{} contains the origin location as the coordinates of the string in the model file. %TODO: implement origin for w-model

<todo: enumerate elements -- after refactoring>

@@

The Z-abstraction level models the the logical formulas (definitions and assertions) that are put on the assertion stack of the SMT-solver.
Generally, a \zformula{} represents the S-expression-based syntax of the SMT-LIB language~\cite{smt-lib}. % the semantics of SMT-LIB formulas defined in~\cite[Chapter 5]{smt-lib}.
Currently, the \zformula{} supports definitions and assertions over \textit{variables} and binary and unary \textit{expressions}.
However, in future it may be extended to support \textit{function symbols} and \textit{binders} (existential and universal quantification, pattern matching and functional type construction).

All expressions of a \zformula{} are \textit{typed} (or \textit{sorted}, in terms of SMT-LIB standard).
Basically, we distinguish \textit{boolean} and \textit{numeric} (integer, bitvector, real) expressions.
The typisation of a \zformula{} is necessary for checking the basic validity of its expressions and converting it to a typed SMT-LIB formula.
For the sake of transparency, each element of a \zformula{} contains the origin location as the reference to the X-element that produced current Z-element. %TODO: implement origin for z-formula

<todo: enumerate elements -- after completing>

@@

Construction of immutable data structures is performed by units that follow the \textit{builder} pattern.
A builder is a mutable object that contains methods "filling" it by the elements that will constitute the result complex object.
The methods contain basic validity checks of the elements.
Once a builder has been built, it cannot be modified.

@@

The input memory model language \cat{} is discussed in Section~\ref{ch:wmm:cat}.
The \tool{ANTLR} grammar for the \cat{} language was extracted from the parser used by \tool{herd} tool%
%
\footnote{\tool{herd} project repository: \url{https://github.com/herd/herdtools7}} %
%
written in OCaml.
%TODO: upload the grammar to github
%The input memory model 
%TODO: add reference to the grammar file in github repository OR include a short version as an appendix}

@@

%TODO: <not implemented yet>
The \wmodel{} is constructed by the converter \texttt{Cat2WmodelConverter}, which calls \texttt{Cat2WmodelConverterVisitor}, the stateless visitor of memory-model ANTLR syntax tree. %TODO: Implement this!!!
Currently, the visitor supports only the declarative subset of the \cat{} language, and does not process the functional-style definitions (it can to be supported by implementing the

@@

The \xgraph{} abstract machine model disposes infinitely many memory units, both local and shared (see Section~\ref{ch:impl:model:xgraph:mem}).
The X-compiler accesses all memory units via the \texttt{XMemoryManager}, which by the end of pre-compilation stage is already initialised (has registered all shared memory units).
However, the memory manager is a stateful component of the compiler as it offers the methods for declaring and removing local memory units dynamically at the compilation time.
At each time of the compilation process, the \texttt{XMemoryManager} can resolve the memory unit by its name.
Following the C standard, local memory units have higher priority over global ones.
Since the C language allows usage of variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} carries the stack of block block-contexts for local variables. %TODO: refactor XMemoryManager!

@@

The invocation hooking module serves as a knowledge base that stores the semantics of functions.
%TODO: rename HookManager -> XHookManager
Once the X-compiler meets the function invocation, it calls the \texttt{XHookManager}, which tries to match the function signature (in case of program in C language -- only the function name) across all signatures it stores.
Once the signature matches, the hook manager intercepts the compilation process and interprets the method according to the algorithm it stores.
%TODO: say sth about the mode
The hook manager operates multiple invocation hooks depending on the user-defined mode.
All invocation hooks implement the interface \texttt{XInvocationHook}.
Current version of \porthos[2] contains two invocation hooks, the \texttt{} and \texttt{XKernelInvocationHook} that contains ...............................

%TODO: rename InvocationHook -> XInvocationHook
%, having access to the X-interpreter (see Section~\ref{ch:impl:proc:x-compiler:compilation}),
